Towards the near-infrared detection of brown dwarf companions: Exploring methods to detect low mass stellar companions from blended spectra \thanks{Based on observations collected at the European Organisation for Astronomical Research in the Southern Hemisphere under ESO programme 089.C-0977(A)}

### 2018-08-17 nir_companions.tex
percent and spaceing

% Abstract of the paper
\begin{abstract}
   {\rd{} In this paper we attempt to directly detect the near-infrared spectrum of candidate brown dwarf (BD) companions around FGK stars to assert or discard their stellar nature. We explore two different methods to probe the faint spectra of BD companions. The first technique involves the direct subtraction of two observations shifted to mutually cancel out the spectra of the host star. The second technique applies a \(\chi^{2}\) fit to the individual observations of a synthetic binary model comprised of two PHOENIX-ACES spectra. The observed spectra are wavelength calibrated and corrected for the atmospheric absorption with the aid of synthetic telluric models. The direct subtraction method failed to detect the signature of the companions due to poor observational constraints (the observations were insufficiently separated in radial velocity). The \(\chi^{2}\) technique developed here is able to fit the component for the host star but unable to successfully detect the faint spectra of even the largest companion in our sample. We explore how the companion recovery fitting performs on simulated observations and discuss reasons for the non-detection observed. From the injection-recovery analysis, this technique in its current form, is insufficient to recover a companion below 3800~K which corresponds to an upper mass limit for the companions around \(\rm 600~M_{Jup}\). This work highlights the challenges in the spectral detection of faint companions. We explore the limitations of the direct subtraction method in the case of small RV separation, and companion detection with synthetic models at low companion/host flux ratios below 1\%, made difficult by the discrepancy between observed and synthetic spectra around  2100~nm.}
\end{abstract}

@@

{\rd{} The search and detection of faint secondary spectra is not only relevant to planetary atmospheres.~\citet{kolbl_detection_2015} developed a method to detect the presence of optical secondary spectra down to a flux ratio of 1\% in the hosts of \emph{Kepler} transit candidates. The presence of which can cause ambiguities in the system configuration, and increase the uncertainty of the measured planet radius. The characterization of the false positive probability rate for Kepler has been found to be as high as  \(\sim\)35\%~\citet{santerne_sophie_2012}.}

@@

At this point one would normally combine all the optimally extracted nod spectra together to improve the signal-to.noise. {\rd{} However, we discovered extended artefacts in the optimally reduced nod spectra. We were not able to identify why these artefacts are large and extended but they occur near the presence of a large cosmic rays or bad pixel spike seen in the rectangular extraction. These artefacts were not observed in previous works using this pipeline in the \textit{H}-band, so there may be a wavelength dependent affect. The variance weighting procedure during the optimal extraction is somehow being heavily affected by these spikes.} An example of this can be seen in Fig.~\ref{fig:nod_artefacts} where a spike in the rectangular extraction corresponds to the large extended features in the optimal extraction. The amplitude of these artefacts created flux deviations in the combined optimally extracted spectra up to \(\sim\)0.5\%.
Therefore, we took measures to remove these artefacts before combining the nod spectra as we are trying to recover companion spectra with expected flux ratios \(\rm {F_2}/{F_1} < 1\% \). {\rd{} Parameters of the pipeline were adjusted to try and remove these artefacts, such as the $\sigma$ rejection limits and the aperture width with limited success. In a small number of instances allowing the aperture width to be automatically adjusted removed the artefacts.}

All nod spectra for each observation were visually inspected together, as shown in  Fig.~\ref{fig:nod_artefacts}, and any spectra containing artefacts were marked. The optimally extracted nods (top panel)  that were identified were replaced with their rectangularly extracted counter-parts (middle panel). An iterative 4-\(\sigma \) rejection algorithm\footnote{Found at \url{https://github.com/jason-neal/nod_combination}} was applied to the replacement rectangular extractions to remove the erroneous pixels that created the artefacts. The \(\sigma\) value for each pixel was calculated as the standard deviation of the nearest 2 pixels on either side of all 8 nod spectra. The rejected pixels were replaced using linear interpolation along the dispersion axis. {\rd{} Out of the 544 nod spectra from individual detectors, 79 (14\%) contained artefacts and were replaced using this technique.}

@@

Much like issues with Th-Ar calibrations this method only works well when there is sufficient coverage of telluric lines on the detector. For the wavelength setting of these observations, the spectra from the second detector (top right panel of Fig.~\ref{fig:detector4allspectra}) only has two large telluric lines present with several small lines, with relative depths smaller than 1\%, which are difficult to identify. This deteriorated the calibration stability for the second detector. With the lack of telluric contamination and stellar lines on the second detector it may have been ideal for the detection of a faint secondary spectra; unfortunately the wavelength calibration quality varies in an inverse way.

@@

\subsection{Wavelength masking}
We apply several wavelength masks to remove wavelength regions from which we cannot extract information.
Firstly, regions near the edges of each detector where the wavelength solution is extrapolated outside of the calibrating telluric lines are removed, reducing the effective size of each detector by about \(10\%\) or \(\sim\)100~pixels.

Secondly, we mask out any remaining artefacts present in the spectra and the centres of deep telluric lines where telluric correction was not corrected properly, sometimes resulting in ``emission-like'' peaks in the corrected spectrum. These factors combined result in masking out around a further 10\% of the observed spectra.

In Sect.~\ref{sec:results} we also apply a further wavelength restriction to mask out regions where there is a large mismatch between the observed spectrum and the closest synthetic spectra to the host. This significantly restricts the wavelength span utilized for that purpose to around only 43\% and the masked regions are visible in Fig.~\ref{fig:visualinspection-hd2118471}.

@@

The combined spectra is continuum normalized by dividing by an exponential fitted to the continuum of the combined spectrum, as we assume we are in the Rayleigh-Jeans regime. This assumption here is wavelength dependent and other appropriate continuum normalization techniques are also valid. In the case of a BD companion around an FGK star investigated here, the continuum is dominated by the contribution from the host star as it contributes the majority of the spectrum with flux ratios around 2110--2160~nm below \(\sim\)1\%.

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.8\hsize]{images/visualize_result_residuals.pdf}
    \caption{Comparison between the observed {HD 211847} spectrum (blue) and the best fit synthetic binary model (orange dashed) for each detector. The bottom section of each panel shows the residuals between the parts of the observation used in the \(\chi^2\) fit and recovered binary model (\(\rm O-C^2\)) in purple. The red dashed line shows the difference between the recovered binary model and the binary model with the exact same parameters except for the estimated companion temperature of 3200~K (\(\rm C^2[3200K]- C^2\)). The grey shading indicated the wavelength regions where masking has been applied. The thinner masked regions that match with cuts in the observed spectra are where the centres of deep (>5\%) telluric lines that have been masked out are.}
    \label{fig:visualinspection-hd2118471}
\end{figure*}

@@

Using the temperature cut-off values, we derive an upper mass limit for the companions around our stars using the~\citet{baraffe_new_2015} evolutionary models, finding the closest point matching the spectral temperature cut-off and \(\rm logg=5.0\). These values are given in Table~\ref{tab:mass_limits} and are between 560 and 618 \(\rm M_{Jup}\). The flux ratio between the cut-off and the host star are also provided for, being between 5 and 15\% in this wavelength span.

@@

For example, with the synthetic model for the companion of {HD 211847}, the average contributions of lines >5\% become only 0.3\% deep in a binary with the Sun-like spectrum. For a companion with a temperature of 2300~K (the lower PHOENIX-ACES temperature limit) the deepest lines contribute lines around 0.1\%.

@@

Our non-detection of binary companions with low flux ratios is consistent with results from other works. For example~\citet{nemravova_xtauri_2016} performed extensive spectral analysis of a quadruple-star system {$\xi$ Tauri} using 227 spectra in 3 different wavelength bands. Of the four stars in the system they were unable to detect the spectral component of the one which had a luminosity ratio below 1\%. {\rd{} The secondary detection in optical spectra using spectral matching of KOI was also only able to reach flux ratios of 1\%~\citet{kolbl_detection_2015}.}

@@

    \begin{figure}
        \includegraphics[width=0.95\hsize]{images/differential.pdf}\\
        \caption{ (Top) A reduced CRIRES observation of {HD 30501} (blue) for detector 1 between 2112--2124~nm along with the tapas telluric absorption model ({orange} dashed) used for the wavelength calibration and telluric correction. (Middle) The telluric corrected spectra. (Bottom) ({blue}) Differential spectra for {HD 30501} between observations 1 and 3. ({orange} dashed) Simulated ``perfect'' differential using PHOENIX-ACES spectra with parameters \(T_{\textrm{eff}} = 2500~K \), \(\rm logg=5.0\), and \(\rm [Fe/H]=0.0 \), with the same \(\Delta RV \) as the observations. The shaded regions indicate where the telluric {green} and host star {blue} spectra are > 4\% deep.}
        \label{fig:spectral_example}
    \end{figure}

    We performed the differential analysis for all targets but only show our most favourable case here, {HD 30501}, because it is the second largest companion in our sample at \(\rm 90~M_{Jup}\) and also has the second largest RV separation between observations. The differential spectra recovered for {HD 30501} is shown at the bottom panel of Fig.~\ref{fig:spectral_example}. The presence of deep (\(>4\% \)) stellar and telluric lines in the original spectrum is shaded by the blue and green regions respectively. This indicates that the features of the differential spectrum near these shaded regions are likely due to imperfect telluric correction and host mutual cancellation.
    The mutual cancellation of the stellar host works well for the \(\sim\)40\%  deep line near 2117~nm, being completely removed, but it does not do so well for the smaller \(\sim\)10\% deep line around 2121.5~nm. The residual for the large \(\sim\)40\% deep telluric line near 2118.5~nm is quite prominent. There is also a wider residual due to three neighbouring lines \(\sim\)10\% deep around 2120~nm which cause features in the differential spectrum. One possible explanation is that the continuum normalization near 2120~nm was influenced by this grouping of lines.

@@

    The two solid vertical lines in Fig.~\ref{fig:diff_amp} indicate the estimated \(\Delta \textrm{RV}\)=1.346\kmps{} separation for our best target, {HD 30501} from Table~\ref{tab:estimatedparameters}, given known orbital parameters and the observation times. This shows that our differentials have severely reduced amplitude, \(<20\%\) relative to well separated individual lines. As the companion spectra are faint and in combination with a host star at 1\% flux ratio the >80\% extra reduction in signal amplitude makes this detection impossible with these observations.

>>>>

% Abstract of the paper
\begin{abstract}
   {\rd{} In this paper we attempt to directly detect the near-infrared spectrum of candidate brown dwarf (BD) companions around FGK stars to assert or discard their stellar nature. We explore two different methods to probe the faint spectra of BD companions. The first technique involves the direct subtraction of two observations shifted to mutually cancel out the spectra of the host star. The second technique applies a \(\chi^{2}\) fit to the individual observations of a synthetic binary model comprised of two PHOENIX-ACES spectra. The observed spectra are wavelength calibrated and corrected for the atmospheric absorption with the aid of synthetic telluric models. The direct subtraction method failed to detect the signature of the companions due to poor observational constraints (the observations were insufficiently separated in radial velocity). The \(\chi^{2}\) technique developed here is able to fit the component for the host star but unable to successfully detect the faint spectra of even the largest companion in our sample. We explore how the companion recovery fitting performs on simulated observations and discuss reasons for the non-detection observed. From the injection-recovery analysis, this technique in its current form, is insufficient to recover a companion below 3800~K which corresponds to an upper mass limit for the companions around \(\rm 600~M_{Jup}\). This work highlights the challenges in the spectral detection of faint companions. We explore the limitations of the direct subtraction method in the case of small RV separation, and companion detection with synthetic models at low companion/host flux ratios below 1 per cent, made difficult by the discrepancy between observed and synthetic spectra around 2100~nm.}
\end{abstract}

@@

{\rd{} The search and detection of faint secondary spectra is not only relevant to planetary atmospheres.~\citet{kolbl_detection_2015} developed a method to detect the presence of optical secondary spectra down to a flux ratio of 1 per cent in the hosts of \emph{Kepler} transit candidates. The presence of which can cause ambiguities in the system configuration, and increase the uncertainty of the measured planet radius. The characterization of the false positive probability rate for Kepler has been found to be as high as  \(\sim\)35 per cent~\citet{santerne_sophie_2012}.}

@@

At this point one would normally combine all the optimally extracted nod spectra together to improve the signal-to.noise. {\rd{} However, we discovered extended artefacts in the optimally reduced nod spectra. We were not able to identify why these artefacts are large and extended but they occur near the presence of a large cosmic rays or bad pixel spike seen in the rectangular extraction. These artefacts were not observed in previous works using this pipeline in the \textit{H}-band, so there may be a wavelength dependent affect. The variance weighting procedure during the optimal extraction is somehow being heavily affected by these spikes.} An example of this can be seen in Fig.~\ref{fig:nod_artefacts} where a spike in the rectangular extraction corresponds to the large extended features in the optimal extraction. The amplitude of these artefacts created flux deviations in the combined optimally extracted spectra up to \(\sim\)0.5 per cent.
Therefore, we took measures to remove these artefacts before combining the nod spectra as we are trying to recover companion spectra with expected flux ratios \(\rm {F_2}/{F_1} < 1\)per cent. {\rd{} Parameters of the pipeline were adjusted to try and remove these artefacts, such as the $\sigma$ rejection limits and the aperture width with limited success. In a small number of instances allowing the aperture width to be automatically adjusted removed the artefacts.}

All nod spectra for each observation were visually inspected together, as shown in Fig.~\ref{fig:nod_artefacts}, and any spectra containing artefacts were marked. The optimally extracted nods (top panel)  that were identified were replaced with their rectangularly extracted counter-parts (middle panel). An iterative 4-\(\sigma \) rejection algorithm\footnote{Found at \url{https://github.com/jason-neal/nod_combination}} was applied to the replacement rectangular extractions to remove the erroneous pixels that created the artefacts. The \(\sigma\) value for each pixel was calculated as the standard deviation of the nearest 2 pixels on either side of all 8 nod spectra. The rejected pixels were replaced using linear interpolation along the dispersion axis. {\rd{} Out of the 544 nod spectra from individual detectors, 79 (14 per cent) contained artefacts and were replaced using this technique.}

@@

Much like issues with Th-Ar calibrations this method only works well when there is sufficient coverage of telluric lines on the detector. For the wavelength setting of these observations, the spectra from the second detector (top right panel of Fig.~\ref{fig:detector4allspectra}) only has two large telluric lines present with several small lines, with relative depths smaller than 1 per cent, which are difficult to identify. This deteriorated the calibration stability for the second detector. With the lack of telluric contamination and stellar lines on the second detector it may have been ideal for the detection of a faint secondary spectra; unfortunately the wavelength calibration quality varies in an inverse way.

@@

\subsection{Wavelength masking}
We apply several wavelength masks to remove wavelength regions from which we cannot extract information.
Firstly, regions near the edges of each detector where the wavelength solution is extrapolated outside of the calibrating telluric lines are removed, reducing the effective size of each detector by about \(10\) per cent or \(\sim\)100~pixels.

Secondly, we mask out any remaining artefacts present in the spectra and the centres of deep telluric lines where telluric correction was not corrected properly, sometimes resulting in ``emission-like'' peaks in the corrected spectrum. These factors combined result in masking out around a further 10 per cent of the observed spectra.

In Sect.~\ref{sec:results} we also apply a further wavelength restriction to mask out regions where there is a large mismatch between the observed spectrum and the closest synthetic spectra to the host. This significantly restricts the wavelength span utilized for that purpose to around only 43 per cent and the masked regions are visible in Fig.~\ref{fig:visualinspection-hd2118471}.

@@

The combined spectra is continuum normalized by dividing by an exponential fitted to the continuum of the combined spectrum, as we assume we are in the Rayleigh-Jeans regime. This assumption here is wavelength dependent and other appropriate continuum normalization techniques are also valid. In the case of a BD companion around an FGK star investigated here, the continuum is dominated by the contribution from the host star as it contributes the majority of the spectrum with flux ratios around 2110--2160~nm below \(\sim\)1 per cent.

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.8\hsize]{images/visualize_result_residuals.pdf}
    \caption{Comparison between the observed {HD 211847} spectrum (blue) and the best fit synthetic binary model (orange dashed) for each detector. The bottom section of each panel shows the residuals between the parts of the observation used in the \(\chi^2\) fit and recovered binary model (\(\rm O-C^2\)) in purple. The red dashed line shows the difference between the recovered binary model and the binary model with the exact same parameters except for the estimated companion temperature of 3200~K (\(\rm C^2[3200K]- C^2\)). The grey shading indicated the wavelength regions where masking has been applied. The thinner masked regions that match with cuts in the observed spectra are where the centres of deep (>5 per cent) telluric lines that have been masked out are.}
    \label{fig:visualinspection-hd2118471}
\end{figure*}

@@

Using the temperature cut-off values, we derive an upper mass limit for the companions around our stars using the~\citet{baraffe_new_2015} evolutionary models, finding the closest point matching the spectral temperature cut-off and \(\rm logg=5.0\). These values are given in Table~\ref{tab:mass_limits} and are between 560 and 618 \(\rm M_{Jup}\). The flux ratio between the cut-off and the host star are also provided for, being between 5 and 15 per cent in this wavelength span.

@@

For example, with the synthetic model for the companion of {HD 211847}, the average contributions of lines >5 per cent become only 0.3 per cent deep in a binary with the Sun-like spectrum. For a companion with a temperature of 2300~K (the lower PHOENIX-ACES temperature limit) the deepest lines contribute lines around 0.1 per cent.

@@

Our non-detection of binary companions with low flux ratios is consistent with results from other works. For example~\citet{nemravova_xtauri_2016} performed extensive spectral analysis of a quadruple-star system {$\xi$ Tauri} using 227 spectra in 3 different wavelength bands. Of the four stars in the system they were unable to detect the spectral component of the one which had a luminosity ratio below 1 per cent. {\rd{} The secondary detection in optical spectra using spectral matching of KOI was also only able to reach flux ratios of 1 per cent~\citet{kolbl_detection_2015}.}

@@

    \begin{figure}
        \includegraphics[width=0.95\hsize]{images/differential.pdf}\\
        \caption{ (Top) A reduced CRIRES observation of {HD 30501} (blue) for detector 1 between 2112--2124~nm along with the tapas telluric absorption model ({orange} dashed) used for the wavelength calibration and telluric correction. (Middle) The telluric corrected spectra. (Bottom) ({blue}) Differential spectra for {HD 30501} between observations 1 and 3. ({orange} dashed) Simulated ``perfect'' differential using PHOENIX-ACES spectra with parameters \(T_{\textrm{eff}} = 2500~K \), \(\rm logg=5.0\), and \(\rm [Fe/H]=0.0 \), with the same \(\Delta RV \) as the observations. The shaded regions indicate where the telluric {green} and host star {blue} spectra are > 4 per cent deep.}
        \label{fig:spectral_example}
    \end{figure}

    We performed the differential analysis for all targets but only show our most favourable case here, {HD 30501}, because it is the second largest companion in our sample at \(\rm 90~M_{Jup}\) and also has the second largest RV separation between observations. The differential spectra recovered for {HD 30501} is shown at the bottom panel of Fig.~\ref{fig:spectral_example}. The presence of deep (\(>4\) per cent% \)) stellar and telluric lines in the original spectrum is shaded by the blue and green regions respectively. This indicates that the features of the differential spectrum near these shaded regions are likely due to imperfect telluric correction and host mutual cancellation.
    The mutual cancellation of the stellar host works well for the \(\sim\)40 per cent deep line near 2117~nm, being completely removed, but it does not do so well for the smaller \(\sim\)10 per cent deep line around 2121.5~nm. The residual for the large \(\sim\)40 per cent deep telluric line near 2118.5~nm is quite prominent. There is also a wider residual due to three neighbouring lines \(\sim\)10 per cent deep around 2120~nm which cause features in the differential spectrum. One possible explanation is that the continuum normalization near 2120~nm was influenced by this grouping of lines.

@@

    The two solid vertical lines in Fig.~\ref{fig:diff_amp} indicate the estimated \(\Delta \textrm{RV}\)=1.346\kmps{} separation for our best target, {HD 30501} from Table~\ref{tab:estimatedparameters}, given known orbital parameters and the observation times. This shows that our differentials have severely reduced amplitude, \(<20\)per cent%\) relative to well separated individual lines. As the companion spectra are faint and in combination with a host star at 1 per cent flux ratio the >80 per cent extra reduction in signal amplitude makes this detection impossible with these observations.
